---
title: "SURV_617_Project"
author: "Rachael Jackson & Sofi Sinozich"
date: "11/17/2020"
output: html_document
---

Load data
```{r}
exam_data <- readRDS("data/exam.rds")
```

Load libraries
```{r}
library(tidyverse)
```

#Create bar chart of health status by examiner.

We observe a lot of variation in distribution of health status scores by examiner.

* Note:
1 = Excellent
2 = Very good
3 = Good
4 = Fair
5 = Poor
8 = Blank but applicable
```{r}
# Distribution of scores given by examiner
exam_data %>% select(PEPTECH,PEP13A) %>%
  filter(PEPTECH <88888 & PEP13A < 8 ) %>% 
  ggplot +
  geom_bar(aes(x = PEP13A)) +
  facet_wrap(PEPTECH ~.)
```

#Create bar chart of referral status by examiner.

We observe some variation in distribution of referral status by examiner, but the vast majority of respondents did not receive referrals that required quick or urgent attention.

* Note: 
Level I = Major medical emergencies that warranted immediate attention by a health care provider
Level II = Major medical findings that warranted attention by a health care provider within 2 weeks
Level III = All other findings reported to the examinee
```{r}
exam_data %>% select(PEPTECH,PEPLEVEL) %>%
  filter(PEPTECH <88888 & PEPLEVEL < 8 ) %>% 
  ggplot +
  geom_bar(aes(x = PEPLEVEL)) +
  facet_wrap(PEPTECH ~.)
```

#Check for confounding--do all proposed variables differ by both health_status and examiner_id?

#Race
Race varies moderately by health status and quite a bit by examiner.
```{r}
#Health status x race
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(race))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x race
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(race))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Gender
Gender varies minimally by health status but moderately by examiner.
```{r}
#Health status x race
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(gender))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x race
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(gender))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Age
Age varies significantly by health status and moderately by examiner.
```{r}
#Health status x age
ggplot(full_data, aes(x=as.factor(health_status), y=age)) + 
  geom_boxplot() 

#Examiner x age
ggplot(full_data, aes(x=as.factor(examiner_id), y=age)) + 
  geom_boxplot()
```

#Region
Region varies a little bit by health status and quite a bit by examiner.
```{r}
#Health status x region
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(region))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x region
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(region))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Locomotion
Locomotion appears to vary significantly by health status and moderately by examiner.
```{r}
#Health status x locomotion
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(locomotion))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x locomotion
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(locomotion))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Paralysis
Paralysis varies a lot by health status and a little by examiner.
```{r}
#Health status x paralysis
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(paralysis))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x paralysis
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(paralysis))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Systolic BP
Systolic BP varies by health status but only varies slightly by examiner.
```{r}
#Health status x systolic_bp
ggplot(full_data, aes(x=as.factor(health_status), y=systolic_bp)) + 
  geom_boxplot() 

#Examiner x systolic_bp
ggplot(full_data, aes(x=as.factor(examiner_id), y=systolic_bp)) + 
  geom_boxplot()
```

#Diastolic BP
Diastolic BP does not appear to have much variation by health status or examiner.
* Note: Systolic and diastolic BP work together to assess hypertension. I recommend we either use systolic BP only (because it shows more variation) or combine them to create a binomial hypertension variable.
```{r}
#Health status x diastolic_bp
ggplot(full_data, aes(x=as.factor(health_status), y=diastolic_bp)) + 
  geom_boxplot() 

#Examiner x diastolic_bp
ggplot(full_data, aes(x=as.factor(examiner_id), y=diastolic_bp)) + 
  geom_boxplot()
```

#Dermatitis
Dermatitis varies a little by health status and moderately by examiner.
* Note: Dermatitis almost appears to be negatively correlated with health status, which tells me it's not an important enough health issue to affect a doctor's health status assessment. I recommend removing this variable from the model
```{r}
#Health status x dermatitis
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(dermatitis))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x dermatitis
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(dermatitis))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#Impairment
Impairment varies significantly by health status and minimally by examiner.
```{r}
#Health status x impairment
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(impairment))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x impairment
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(impairment))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```
#Infection
Infection varies quite a bit by both health status and examiner.
```{r}
#Health status x infection
ggplot(full_data, aes(x=as.factor(health_status), fill=as.factor(infection))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")

#Examiner x infection
ggplot(full_data, aes(x=as.factor(examiner_id), fill=as.factor(infection))) + 
  geom_bar(position = "fill")+ 
  scale_y_continuous(labels = scales::percent, breaks = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)) +
  ylab("Percent")
```

#BMI
BMI varies moderately by health status but minimally by examiner.
```{r}
#Health status x BMI
ggplot(full_data, aes(x=as.factor(health_status), y=bmi)) + 
  geom_boxplot() 

#Examiner x BMI
ggplot(full_data, aes(x=as.factor(examiner_id), y=bmi)) + 
  geom_boxplot()
```

#RBC Count
RBC count varies minimally by both health status and examiner.
*Note: I did a quick Google check, and it looks like RBC count and hemoglobin measure roughly the same thing--anemia. Neither appears to be strongly correlated with health status or examiner, so we could probably remove both from the model, but should definitely remove one from the model.
```{r}
#Health status x RBC count
ggplot(full_data, aes(x=as.factor(health_status), y=rbc_count)) + 
  geom_boxplot() 

#Examiner x RBC count
ggplot(full_data, aes(x=as.factor(examiner_id), y=rbc_count)) + 
  geom_boxplot()
```

#WBC Count
WBC count varies minimally by both health status and examiner.
*Note: WBC count is primarily an indicator of infection. Since we already have an infection variable, which appears to be more strongly correlated with health status and examiner than wbc_count is, I recommend removing wbc_count from the model.
```{r}
#Health status x WBC count
ggplot(full_data, aes(x=as.factor(health_status), y=wbc_count)) + 
  geom_boxplot() 

#Examiner x WBC count
ggplot(full_data, aes(x=as.factor(examiner_id), y=wbc_count)) + 
  geom_boxplot()
```

#Platelet Count
Platelet count varies minimally by both health status and examiner.
*Note: Platelet count is another indicator of infection, as well as cancer (and some cancer treatments), kidney infections, medications, too much alcohol, or an inherited condition. Since there's not a ton of correlation between plt_count and health status or examiner, and since it overlaps a bit with the infection variable, I recommend removing it from the model.
```{r}
#Health status x Platelet count
ggplot(full_data, aes(x=as.factor(health_status), y=plt_count)) + 
  geom_boxplot() 

#Examiner x Platelet count
ggplot(full_data, aes(x=as.factor(examiner_id), y=plt_count)) + 
  geom_boxplot()
```

#Hemoglobin
Hemoglobin varies minimally by both health status and examiner.
*Note: As noted above, hemoglobin is primarily an indicator of RBC count. (Too little hemoglobin = too few RBCs = anemia) Since it's not strongly correlated with health status or examiner and overlaps with RBC count, in terms of medical meaning, I recommend removing it from the model.
```{r}
#Health status x hemoglobin
ggplot(full_data, aes(x=as.factor(health_status), y=hemoglobin)) + 
  geom_boxplot() 

#Examiner x hemoglobin
ggplot(full_data, aes(x=as.factor(examiner_id), y=hemoglobin)) + 
  geom_boxplot()
```






















